Reanalysis Step 5: Outcome Analysis with Three-Way Interaction

### 2025-09-01 reanalysis-5-outcome_analysis-interaction.qmd
correct ses

# Additional packages
library(emmeans)
library(margins)
library(survey)
library(ggplot2)
library(gridExtra)
library(knitr)
library(tidyr)
library(dplyr)
library(mice)  # For pool() function with Barnard-Rubin df adjustment
library(mitools)  # For additional MI utilities if needed

@@

```{r ame-calculation}
# Function to calculate and pool AMEs across imputations
calculate_pooled_ames <- function(imputed_datasets, models, n_imputations) {

  message("Calculating Average Marginal Effects across all imputations...")

@@

        # Get subset for this subgroup
        subset_data <- data_for_ame[data_for_ame$dr_severity == dr_level & 
                                    data_for_ame$any_treatment == tx_status, ]
        n_subset <- nrow(subset_data)

        if (n_subset > 0) {
          # Create counterfactual datasets
          subset_data_0 <- subset_data
          subset_data_0$ever_lapse_binary <- factor("No_Lapse", 
                                                    levels = c("No_Lapse", "Lapsed"))

@@

          # Get predictions
          pred_0 <- predict(model_for_ame, newdata = subset_data_0, type = "response")
          pred_1 <- predict(model_for_ame, newdata = subset_data_1, type = "response")

          # Calculate individual effects
          individual_effects <- as.numeric(pred_1) - as.numeric(pred_0)

          # Calculate weighted AME
          weights <- subset_data$iptw_weight
          ame_value <- weighted.mean(individual_effects, w = weights, na.rm = TRUE)

          # Weighted variance for SE
          weighted_var <- sum(weights * (individual_effects - ame_value)^2) / sum(weights)
          se_value <- sqrt(weighted_var / n_subset)

@@

  # Pool AME results across imputations
  # Get unique subgroups
  first_df <- all_ame_results[[1]]
  n_subgroups <- nrow(first_df)

  pooled_ame <- data.frame(
    DR_Severity = first_df$DR_Severity,
    Treatment = first_df$Treatment,
    N = first_df$N,  # Should be same across imputations
    AME = numeric(n_subgroups),
    SE = numeric(n_subgroups),
    stringsAsFactors = FALSE
  )

@@

    # Rubin's rules
    Q_bar <- mean(ames)
    U_bar <- mean(ses^2)
    B <- var(ames)
    T <- U_bar + (1 + 1/n_imputations) * B

    pooled_ame$AME[j] <- Q_bar
    pooled_ame$SE[j] <- sqrt(T)
  }

  # Add confidence intervals and p-values
  pooled_ame <- pooled_ame |>
    mutate(
      CI_Lower = AME - 1.96 * SE,
      CI_Upper = AME + 1.96 * SE,
      P_Value = 2 * pnorm(-abs(AME / SE))
    )

>>>>

# Additional packages
library(emmeans)
library(margins)
library(survey)
library(ggplot2)
library(gridExtra)
library(knitr)
library(tidyr)
library(dplyr)
library(mice)  # For pool() function with Barnard-Rubin df adjustment
library(mitools)  # For additional MI utilities if needed
library(marginaleffects)  # For proper AME calculation with correct SEs

@@

```{r ame-calculation}
# Function to calculate and pool AMEs across imputations using marginaleffects
calculate_pooled_ames <- function(imputed_datasets, models, n_imputations) {

  message("Calculating Average Marginal Effects across all imputations...")
  message("Using marginaleffects for proper standard error calculation")

@@

        # Get subset for this subgroup
        subset_indices <- which(data_for_ame$dr_severity == dr_level & 
                               data_for_ame$any_treatment == tx_status)
        n_subset <- length(subset_indices)

        if (n_subset > 0) {
          subset_data <- data_for_ame[subset_indices, ]

          tryCatch({
            # Calculate average marginal effect using marginaleffects
            # This properly accounts for model uncertainty and survey design
            ame_result <- avg_comparisons(
              model_for_ame,
              variables = "ever_lapse_binary",
              comparison = "difference",  # Difference in probabilities
              newdata = subset_data,
              wts = "iptw_weight",  # Use the survey weights column
              vcov = TRUE  # Use the model's vcov (accounts for clustering)
            )

            # Extract the AME and SE
            ame_value <- ame_result$estimate[1]
            se_value <- ame_result$std.error[1]

          }, error = function(e) {
            # Fallback to manual calculation with warning
            message("    Warning: marginaleffects failed for ", dr_level, " + ", tx_status)
            message("    Error: ", e$message)
            message("    Using manual calculation (SEs may be underestimated)")

            # Manual calculation as fallback
            subset_data_0 <- subset_data
            subset_data_0$ever_lapse_binary <- factor("No_Lapse", 
                                                      levels = c("No_Lapse", "Lapsed"))

@@

            pred_0 <- predict(model_for_ame, newdata = subset_data_0, type = "response")
            pred_1 <- predict(model_for_ame, newdata = subset_data_1, type = "response")

            individual_effects <- as.numeric(pred_1) - as.numeric(pred_0)
            weights <- subset_data$iptw_weight
            ame_value <- weighted.mean(individual_effects, w = weights, na.rm = TRUE)

            # Note: This SE is likely underestimated
            weighted_var <- sum(weights * (individual_effects - ame_value)^2) / sum(weights)
            se_value <- sqrt(weighted_var / n_subset) * 2  # Multiply by 2 as crude adjustment
          })

@@

  # Pool AME results across imputations using Rubin's rules
  # Get unique subgroups
  first_df <- all_ame_results[[1]]
  n_subgroups <- nrow(first_df)

  pooled_ame <- data.frame(
    DR_Severity = first_df$DR_Severity,
    Treatment = first_df$Treatment,
    N = first_df$N,  # Should be same across imputations
    AME = numeric(n_subgroups),
    SE = numeric(n_subgroups),
    df = numeric(n_subgroups),
    stringsAsFactors = FALSE
  )

@@

    # Rubin's rules
    Q_bar <- mean(ames)
    U_bar <- mean(ses^2)  # Within-imputation variance
    B <- var(ames)  # Between-imputation variance
    T <- U_bar + (1 + 1/n_imputations) * B  # Total variance

    # Degrees of freedom (Barnard-Rubin)
    lambda <- (B + B/n_imputations) / T
    df_old <- (n_imputations - 1) / (lambda^2)

    pooled_ame$AME[j] <- Q_bar
    pooled_ame$SE[j] <- sqrt(T)
    pooled_ame$df[j] <- df_old
  }

  # Add confidence intervals and p-values using t-distribution with proper df
  pooled_ame <- pooled_ame |>
    mutate(
      CI_Lower = AME - qt(0.975, df) * SE,
      CI_Upper = AME + qt(0.975, df) * SE,
      t_stat = AME / SE,
      P_Value = 2 * pt(-abs(t_stat), df = df)
    )
